A Tutorial on Tailored Power Simulations for Experimental Study Designs with Generalized Linear Mixed Models

### 2023-09-11 docs/manuscript.Rmd
anne's comments

<!-- General short intro paragraph, need for well-powered and analyzed experimental research -->
When planning experimental research, it is essential to determine an appropriate sample size [@lakensSampleSizeJustification2022] to ensure that the results obtained are both robust and informative, and to use appropriate statistical models to analyze the data. However, the recent replication crisis in experimental research, such as in Human-Computer Interaction (HCI), <!-- SG: is the HCI needed here? Also, can we claim a general replication crisis in experimental research? For instance, have no idea if experimental physics also has a problem. Maybe "in several disciplines --> has illustrated many challenges surrounding the reproducibility and reliability of findings [@robertsonImprovingStatisticalPractice2016]. As a result, there is a growing need for rigorous statistical methodology and the adoption of well-powered experimental designs. While software solutions exist for simple statistical models and experimental designs, many researchers lack the skills and tools to conduct power analyses for more complex research designs using the flexible generalized linear mixed models (GLMM) framework. In the present work, we provide a tutorial comprised of a concrete example for custom a priori power analyses using data simulations based on GLMMs.

@@

<!-- Some basics on statistical power -->
In empirical research relying on hypothesis testing, the most common strategy for determining an adequate sample size is based on statistical power [@lakensSampleSizeJustification2022]. Statistical power is defined as the probability that a hypothesis test has a significant p-value when analyzing repeated samples from a population with a true effect of some pre-specified size. Less formally, power is also described as the probability that a hypothesis test accepts the alternative hypothesis $H_1$ if $H_1$ is indeed true. If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects with high probability the effects or relationships being investigated <!-- SG: I think there is something wrong with the logic of this sentence. Should it be like this? If the sample size used for data collection is insufficient to detect the effects or relationships being investigated with high probability, the study would be considered "underpowered"-->, this particular study would be considered "underpowered".

<!-- what are the resulting issues with underpowered studies? -->
Conducting underpowered research has many negative consequences. First, relying on underpowered experiments may yield inconclusive (if low power is taken into account) or misleading (if low power is ignored) results, hindering the accumulation of knowledge. Second, underpowered studies waste resources by consuming time, effort, and funding without delivering meaningful results.

@@

<!-- define power analysis -->
A power analysis represents the act of calculating the statistical power for a given true effect and sample size. 
By running the calculation for a range of sample sizes "a priori" (i.e., before data collection), the required sample size can be determined so that researchers find an assumed true effect with the desired statistical power. 
Thereby, power analysis offers a valuable contribution to the research process by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value.

<!-- advantages of an a priori power analysis -->
Conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for analyzing the data and answering their research questions.
Also, while conducting a proper power analysis, one has to consider every aspect of the experimental design and will possibly notice statistical or design challenges before starting with data collection. 
This gives researchers a great opportunity to improve the design. Especially in fast-moving research fields, adding a solid sample size calculation to the process can act as a safeguard for ensuring high-quality research.

<!-- external incentives to do a power analysis -->
Moreover, many journals and funding agencies now require that a power analysis is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings. By incorporating power analysis into research planning, researchers can enhance the replicability and credibility of their work, contributing to the advancement of the respective research field through more robust insights. <!-- FP: this last sentence seems like a repetition. wouldn't some sentence better fit here stating that proficiency with power analysis improves your career chances in academia by increasing the probability to get papers published and grants funded (which is a new argument we haven't already made elsewhere); TK: I like this argument. Might be a little idealistic though since bc flashy results get you ahead mostly :D   -->

<!-- standard software is available for "simple" models and study designs -->
For simple statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups), user-friendly software for power analysis is readily available [@faulStatisticalPowerAnalyses2009]. <!-- FP: we could reintroduce the old section her on: with more complex models and designs, power analysis is more complicated because no analytic solution is possible, power has to be estimated by simulation, and there is no user-friendly software that includes all relevant designs.; TK: yes let's do this. I am not entirely sure what you mean with "analytic" solution, i think we should define this. SG: I am not sure what the plan is here. This sentence alone makes little sense. The argument that these tools cannot be used or adjusted for more complex models is missing. If you want to link it to the next section, it would argue that these simple statistical methods are often not sufficient for researchers to answer their research questions and that more complex designs are required. I would not go in detail here, that no tools for GLMM exist, since this argument is already mentioned below-->

@@

<!-- Advantages of GLMMs over "simpler" models -->
As study designs become more complex, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022]. GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data from experimental research, such as in HCI, because they offer a flexible framework for analyzing data with non-normal and categorical outcomes while accounting for both fixed and random effects [@kapteinUsingGeneralizedLinear2016; @fahrmeirRegressionModelsMethods2021].

<!-- intro into GLMMs -->
GLMMs are an extension of LMMs (Linear Mixed Models), which are, in turn, extensions of classic linear regression models that account for correlated data and hierarchical structures [@fahrmeirRegressionModelsMethods2021]. In this context, correlated data means that observations within a given dataset are not independent of each other. In other words, the values of the dependent variable for one observation may be related to the values of the dependent variable for other observations in a systematic way. This correlation can arise for various reasons, such as repeated measurements on the same subjects over time, observations clustered within certain groups or locations, or data collected at different levels of granularity. Hierarchical structures can be found in datasets that have a nested or hierarchical organization. For example, one might have data on students nested within schools, employees nested within departments, or patients nested within hospitals. In such cases, the observations within each higher-level grouping (e.g., school, department, hospital) tend to be more similar to each other than to observations in other groups.
LMMs are used when the outcome variable is continuous and follows a normal distribution (when conditioned on the predictor variables). They allow for the modelling of fixed effects, which capture the relationships between predictors and the outcome, as well as random effects, which account for the correlation and variability within groups or subjects. Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across the groups or subjects. 
As mentioned, GLMMs extend the LMM framework to accommodate non-normal and categorical outcome variables. They are used when the outcome variable does not follow a normal distribution but instead belongs to a different distribution family (e.g., binomial) <!-- SG: do you think examples for the different types of variables are needed-->. GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable. The link function allows for modelling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the response variable.

@@

<!-- what are the existing options for a priori power analysis for GLMMs / where are existing tutorials limited -->
Currently available software packages for power simulations <!-- FP: because you removed the sentence on the difference between analytic and simulation based power analysis, it is now not clear why from now on we always talk about power simulation; TK: sorry if I have accidentally cut this distinction when merging versions, i think it is super important to have the definition between the two here.  --> do not include GLMMs, making it necessary to build data simulations tailored specifically to the study design [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022]. 
A number of tutorials have been published describing how to perform such power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021]. 
However, most of these tutorials only cover linear mixed models (LMMs) and focus on the most common designs (e.g., INSERT EXAMPLE). This narrow focus provides little guidance for researchers faced with more complex study designs (e.g., INSERT EXAMPLE), which are very common [@brownIntroductionLinearMixedEffects2021]. 
The necessary assumptions for power simulations in GLMMs include assumptions about the distributional form of the outcome variable, the random effects, and the error structure. The distributional assumption specifies the family of distributions for the outcome variable, such as Gaussian, Poisson, or binomial. Assumptions about the random effects include the assumption of normality and the covariance structure among the random effects. Additionally, assumptions about the error structure, such as independence or correlation, must be specified. Interpreting these assumptions entails understanding the underlying assumptions of the model and ensuring they align with the characteristics of the data being analyzed.
Here, existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters. However, in practice, knowledge about those parameters from prior studies is often limited, specifying assumptions a practical challenge [@maxwellSampleSizePlanning2008].

<!-- Summary / Wrap-Up, pick up main arguments and what the paper provides, keep this short! -->
Based on the need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a data simulation and an a priori power simulation for a more complex study design. Thereby, we aim to equip researchers with the tools needed to simulate their own data and determine appropriate sample sizes for their own research.

@@

<!-- describe the lme4 package and how it works-->
In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015] which is a state-of-the-art tool for fitting frequentist GLMMs.[^1]

@@

## Our specific GLMM

<!-- deep dive into our specific example -->
In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021]. $$
g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta <!-- SG: Sorry for maybe a stupid question, but what does E stand for? and why are model parameters β mentioned but not in the equation?-->
$$ Equivalently, the conditional expected value is modelled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$. $$
E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = g^{-1}(\eta)
$$ If the dependent variable (i.e., diagnostic decision)<!-- SG: not sure if diagnostic accuracy as a term is useful as a binary dependent variable since it is usually calculated as a percentage over many cases.)  --> $Y$ is a binary variable with values $0$ (i.e. inaccurate), or $1$ (i.e., accurate), the conditional expected value is equivalent to the probability: $$
P_{si} := P(Y = 1|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})
$$ In our case study, $P_{si}$ is the conditional probability that a subject $s$ gives the correct response to item (i.e., CT scan) $i$.

@@

```{r}
simulate <- function(n_subjects = 100, n_items = 50,
  b_0 = 0.847, b_e = 1.350, b_a = -1.253, b_c = 2.603,
  b_ea = 0.790, b_ec = -1.393,
  sd_u0s = 0.5, sd_u0i = 0.5, ...){
  require(dplyr)
  require(faux)
  # simulate design
  dat <- add_random(subject = n_subjects, item = n_items) %>%
    add_between("subject", expert = c(1, 0), .prob = c(0.25, 0.75)) %>%
    mutate(advice_present = rbinom(n(), 1, prob = 2/3)) %>%
    mutate(advice_correct = if_else(advice_present == 1, 
                                    rbinom(n(), 1, prob = 0.8), 0)) %>%
    # add random effects
    add_ranef("subject", u0s = sd_u0s) %>%
    add_ranef("item", u0i = sd_u0i) %>%
    # compute dependent variable
    mutate(linpred = b_0 + u0i + u0s +
        b_e * expert + b_a * advice_present + b_c * advice_correct +
        b_ea * expert * advice_present + b_ec * expert * advice_correct) %>%
    mutate(y_prob = plogis(linpred)) %>%
    mutate(y_bin = rbinom(n = n(), size = 1, prob = y_prob))
  dat
}
```

>>>>

<!-- General short intro paragraph, need for well-powered and analyzed experimental research -->
When planning experimental research, it is essential to determine an appropriate sample size [@lakensSampleSizeJustification2022] to ensure that the results obtained are both robust and informative, and to use appropriate statistical models to analyze the data. However, the recent replication crisis in experimental research, such as in Human-Computer Interaction (HCI), <!-- SG: is the HCI needed here? Also, can we claim a general replication crisis in experimental research? For instance, have no idea if experimental physics also has a problem. Maybe "in several disciplines. AK: I agree with Susanne. I am not sure whether the HCI is needed. Seems a bit arbitrary. Why does this apply to HCI in particular? Either work with an example or keep more generel --> has illustrated many challenges surrounding the reproducibility and reliability of findings [@robertsonImprovingStatisticalPractice2016]. As a result, there is a growing need for rigorous statistical methodology and the adoption of well-powered experimental designs. While software solutions exist for simple statistical models and experimental designs, many researchers lack the skills and tools to conduct power analyses for more complex research designs using the flexible generalized linear mixed models (GLMM) framework. In the present work, we provide a tutorial comprised of a concrete example for custom a priori power analyses using data simulations based on GLMMs.
<!-- AK: Great, I like the Intro. I take my previous comment back - I think this can stay as it is now, without examples. I feel that it is not tailored towards HCI, but I also don't think that's a big issue.-->

@@

<!-- Some basics on statistical power -->
In empirical research relying on hypothesis testing, the most common strategy for determining an adequate sample size is based on statistical power [@lakensSampleSizeJustification2022]. Statistical power is defined as the probability that a hypothesis test has a significant p-value when analyzing repeated samples from a population with a true effect of some pre-specified size. Less formally, power is described as the probability that a hypothesis test accepts the alternative hypothesis $H_1$ if $H_1$ is indeed true.<!-- AK: Just a very minor thing: I learned that power is the probability that a statistical test correctly rejects the null hypothesis when the alternative hypothesis is true. That is basically the same as what you wrote, but maybe a bit clearer because you don't have H1 and H1 right after another. But up to you! --> If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to detect the effects with high probability the effects or relationships being investigated <!-- SG: I think there is something wrong with the logic of this sentence. Should it be like this? If the sample size used for data collection is insufficient to detect the effects or relationships being investigated with high probability, the study would be considered "underpowered"-->, this particular study would be considered "underpowered".<!--AK: I agree with Susanne. This is not a sentence anymore. -->

<!-- what are the resulting issues with underpowered studies? -->
Conducting underpowered research has many negative consequences. First, relying on underpowered experiments may yield inconclusive (if low power is taken into account<!--AK: Unclear what this means in this context-  how is it "taken into account"? -->) or misleading (if low power is ignored) results, hindering the accumulation of knowledge. Second, underpowered studies waste resources by consuming time, effort, and funding without delivering meaningful results.

@@

<!-- define power analysis -->
A power analysis represents the act of calculating the statistical power for a given true effect and sample size. 
By running the calculation for a range of sample sizes "a priori" (i.e., before data collection), the required sample size can be determined so that researchers find an assumed true effect with the desired statistical power. <!--AK: I am not sure, but isn't this the definition of power simulation more or less? Maybe you could introduce that here already. -->
Thereby, power analysis offers a valuable contribution to the research process by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value.<!--AK: This is already an advantage. I would combine this with the part below.-->

<!-- advantages of an a priori power analysis -->
Conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for analyzing the data and answering their research questions.
Also, when conducting a proper power analysis, researchers have to consider every aspect of the experimental design and will notice statistical or design challenges before starting with data collection. 
Especially in fast-moving research fields<!--AK: I don't understand how this applies to fast-moving research fields (btw, what is that??) in particular. I would delere this.-->, adding a solid sample size calculation to the process can act as a safeguard for ensuring high-quality research.

<!-- external incentives to do a power analysis -->
Moreover, many journals and funding agencies now require that a power analysis is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings. By incorporating power analysis into research planning, researchers can enhance the replicability and credibility of their work, contributing to the advancement of the respective research field through more robust insights. <!-- FP: this last sentence seems like a repetition. wouldn't some sentence better fit here stating that proficiency with power analysis improves your career chances in academia by increasing the probability to get papers published and grants funded (which is a new argument we haven't already made elsewhere); TK: I like this argument. Might be a little idealistic though since bc flashy results get you ahead mostly :D  AK: I agree, the first part is repetitive. -->

<!-- standard software is available for "simple" models and study designs -->
For simple statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups), user-friendly software for power analysis is readily available [@faulStatisticalPowerAnalyses2009]. <!-- FP: we could reintroduce the old section her on: with more complex models and designs, power analysis is more complicated because no analytic solution is possible, power has to be estimated by simulation, and there is no user-friendly software that includes all relevant designs.; TK: yes let's do this. I am not entirely sure what you mean with "analytic" solution, i think we should define this. SG: I am not sure what the plan is here. This sentence alone makes little sense. The argument that these tools cannot be used or adjusted for more complex models is missing. If you want to link it to the next section, it would argue that these simple statistical methods are often not sufficient for researchers to answer their research questions and that more complex designs are required. I would not go in detail here, that no tools for GLMM exist, since this argument is already mentioned below; AK: I agree with Susanne.-->

@@

<!-- Advantages of GLMMs over "simpler" models -->
As study designs become more complex<!--AK: I am not even sure whether this is true. Do study designs become more complex or do we become more aware of the pitfalls of underpowered studies? I would say it's the latter.-->, researchers require more sophisticated statistical models to capture the nuanced relationships and hierarchical structures introduced by their study designs [@yarkoniGeneralizabilityCrisis2022]. GLMMs (also called multilevel models) are gaining increasing popularity in analyzing data from experimental research, such as in HCI<!--AK: This "such as in HCI" always sounds super weird. It becomes apparant that you need to mention this for the submission. I would eitehr remove this entirely or explain this more in context of HCI research (using an example from that field)-->, because they offer a flexible framework for analyzing data with non-normal and categorical outcomes while accounting for both fixed and random effects [@kapteinUsingGeneralizedLinear2016; @fahrmeirRegressionModelsMethods2021].

<!-- intro into GLMMs -->
GLMMs are an extension of LMMs (Linear Mixed Models), which are, in turn, extensions of classic linear regression models that account for correlated data and hierarchical structures  [@fahrmeirRegressionModelsMethods2021]. In this context, correlated data means that observations within a given dataset are not independent of each other. In other words, the values of the dependent variable for one observation may be related to the values of the dependent variable for other observations in a systematic way. This correlation can arise for various reasons, such as repeated measurements on the same subjects over time, observations clustered within certain groups or locations, or data collected at different levels of granularity. Hierarchical structures can be found in datasets that have a nested or hierarchical organization. For example, one might have data on students nested within schools, employees nested within departments, or patients nested within hospitals. In such cases, the observations within each higher-level grouping (e.g., school, department, hospital) tend to be more similar to each other than to observations in other groups.
LMMs are used when the outcome variable is continuous and follows a normal distribution (when conditioned on the predictor variables). They allow for the modelling of fixed effects, which capture the relationships between predictors and the outcome, as well as random effects, which account for the correlation and variability within groups or subjects. Random effects are typically assumed to follow a normal distribution with a mean of zero and a variance that quantifies the heterogeneity across the groups or subjects. 
As mentioned, GLMMs extend the LMM framework to accommodate non-normal and categorical outcome variables. They are used when the outcome variable does not follow a normal distribution but instead belongs to a different distribution family (e.g., binomial) <!-- SG: do you think examples for the different types of variables are needed-->. GLMMs incorporate both fixed and random effects, similar to LMMs, but also involve a link function that connects the linear predictor to the expected value of the outcome variable. The link function allows for modelling the relationship between predictors and the outcome in a non-linear way that is appropriate for the specific distribution family of the response variable.<!--AK: And here I would add one research example. Something like, "Imagine an experiment where experts are repeatedly asked for... In this case, the data is not independent because ... I would add an example that explains both aspects you talk about before: 1) The dependency, 2) the outcome distribution (e.g., binomial). I think this would make it much easier for the reader to follow you here.-->

@@

<!-- what are the existing options for a priori power analysis for GLMMs / where are existing tutorials limited -->
Currently available software packages for power simulations <!-- FP: because you removed the sentence on the difference between analytic and simulation based power analysis, it is now not clear why from now on we always talk about power simulation; TK: sorry if I have accidentally cut this distinction when merging versions, i think it is super important to have the definition between the two here.  --> do not include GLMMs, making it necessary to build data simulations tailored specifically to the study design [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022]. 
A number of tutorials have been published describing how to perform such power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021]. 
However, most of these tutorials only cover linear mixed models (LMMs) and focus on the most common designs (e.g., INSERT EXAMPLE). This narrow focus provides little guidance for researchers faced with more complex study designs (e.g., INSERT EXAMPLE), which are very common <!--AK: Are very common how? Unclear sentence-->[@brownIntroductionLinearMixedEffects2021]. 
The necessary assumptions for power simulations in GLMMs include assumptions about the distributional form of the outcome variable, the random effects, and the error structure. The distributional assumption specifies the family of distributions for the outcome variable, such as Gaussian, Poisson, or binomial. Assumptions about the random effects include the assumption of normality<!--AK: Normality of what?--> and the covariance structure among the random effects<!--AK: Unclear what this is-->. Additionally, assumptions about the error structure, such as independence or correlation, must be specified. Interpreting these assumptions entails understanding the underlying assumptions <!--AK: Repeated "assumptions". Find other word.-->of the model and ensuring they align with the characteristics of the data being analyzed.
Existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters. However, in practice, knowledge about those parameters from prior studies is often limited, specifying assumptions <!--AK: Again "assumptions".-->a practical challenge [@maxwellSampleSizePlanning2008].

<!-- Summary / Wrap-Up, pick up main arguments and what the paper provides, keep this short! -->
Based on the need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a data simulation and an a priori power simulation for a more complex study design<!--AK: I would just write GLMM.-->. Thereby, we aim to equip researchers with the tools needed to simulate their own data and determine appropriate sample sizes for their own research.

@@

<!-- describe the lme4 package and how it works-->
In our case study, we use the lme4 R package [@batesFittingLinearMixedEffects2015] which is a state-of-the-art tool for fitting frequentist GLMMs<!--AK: I think you need a ref for this statement.-->.[^1]

@@

## Our specific GLMM
<!-- deep dive into our specific example -->
In a GLMM, the expected value of the dependent variable Y conditioned on the vector of predictor variables $\mathbf{X}$ and random effects $\mathbf{U}$, transformed by a link function $g()$ is modeled as a linear combination $\eta$ of the predictor variables $\mathbf{X}$, the random effects $\mathbf{U}$ and the model parameters $\mathbf{\beta}$ [@fahrmeirRegressionModelsMethods2021]. $$
g(E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = \eta <!-- SG: Sorry for maybe a stupid question, but what does E stand for? and why are model parameters beta mentioned but not in the equation?-->
$$ Equivalently, the conditional expected value is modelled as the linear combination $\eta$, transformed by the inverse link function $g^{-1}()$. $$
E(Y|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})) = g^{-1}(\eta)
$$ If the dependent variable (i.e., diagnostic decision)<!-- SG: not sure if diagnostic accuracy as a term is useful as a binary dependent variable since it is usually calculated as a percentage over many cases.)  --> $Y$ is a binary variable with values $0$ (i.e. inaccurate), or $1$ (i.e., accurate), the conditional expected value is equivalent to the probability: $$
P_{si} := P(Y = 1|\mathbf{X}=\mathbf{x},\mathbf{U}=\mathbf{u})
$$ In our case study, $P_{si}$ is the conditional probability that a subject $s$ gives the correct response to item (i.e., CT scan) $i$.

@@

```{r}
simulate <- function(n_subjects = 100, n_items = 50,
  b_0 = 0.847, b_e = 1.350, b_a = -1.253, b_c = 2.603,
  b_ea = 0.790, b_ec = -1.393,
  sd_u0s = 0.5, sd_u0i = 0.5, ...){
  require(dplyr)
  require(faux)
  # simulate design
  dat <- add_random(subject = n_subjects, item = n_items) %>%
    add_between("subject", expert = c(1, 0), .prob = c(0.25, 0.75)) %>%
    mutate(advice_present = rbinom(n(), 1, prob = 2/3)) %>%
    mutate(advice_correct = if_else(advice_present == 1L, 
                                    rbinom(n(), 1L, prob = 0.8), 0L)) %>%
    # add random effects
    add_ranef("subject", u0s = sd_u0s) %>%
    add_ranef("item", u0i = sd_u0i) %>%
    # compute dependent variable
    mutate(linpred = b_0 + u0i + u0s +
        b_e * expert + b_a * advice_present + b_c * advice_correct +
        b_ea * expert * advice_present + b_ec * expert * advice_correct) %>%
    mutate(y_prob = plogis(linpred)) %>%
    mutate(y_bin = rbinom(n = n(), size = 1, prob = y_prob))
  dat
}
```
